%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                          % paper

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
                                                          \usepackage[utf8]{inputenc}
\usepackage[utf8]{inputenc}
\overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document



% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{fancyhdr}
\setlength{\headheight}{15.2pt}
\usepackage[]{algorithm2e}
\usepackage{url}

\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{defi}{Definition}
\newtheorem{prop}{Proposition}
\newtheorem{conj}{Conjecture}

\DeclareMathOperator*{\ch}{ch}

\title{\LARGE \bf
%modif
On the list coloring of 1-band buffering cellular graphs
}

%\author{ \parbox{3 in}{\centering Huibert Kwakernaak*
%         \thanks{*Use the $\backslash$thanks command to put information here}\\
%         Faculty of Electrical Engineering, Mathematics and Computer Science\\
%         University of Twente\\
%         7500 AE Enschede, The Netherlands\\
%         {\tt\small h.kwakernaak@autsubmit.com}}
%         \hspace*{ 0.5 in}
%         \parbox{3 in}{ \centering Pradeep Misra**
%         \thanks{**The footnote marks may be inserted manually}\\
%        Department of Electrical Engineering \\
%         Wright State University\\
%         Dayton, OH 45435, USA\\
%         {\tt\small pmisra@cs.wright.edu}}
%}

\author{Marine Collery$^{1}$ and Benjámin Martin Seregi$^{2}$% <-this % stops a space
% \thanks{*This work was not supported by any organization}% <-this % stops a space
\thanks{$^{1}$ student of Research Methodology and Scientific Writing Course at KTH Kista P1P22017. e-mail: collery@kth.se}%
\thanks{$^{2}$ student of Research Methodology and Scientific Writing Course at KTH Kista P1P22017. e-mail: seregi@kth.se}%
}

\begin{document}


\maketitle
\thispagestyle{fancy}
\fancyhf{}
\chead{KTH Royal Institute of Technology | II2202, Fall 2017, Period 1-2 | \today}

\begin{abstract}
The optimal channel allocation problem (CAP) in cellular networks is often formulated in a graph theoretic framework. One of its variants\---where each access point knows the list of its free channels\---is related the so-called list coloring problem. In spite of the fact that the list coloring problem is NP-complete for arbitrary graphs, we showed that there exists a polynomial time algorithm for 1-band buffering callular graphs; and as a corollary it turned out the choice number of such a graphs is at most 4. In addition, we carried out performance comparisons between the existing integer linear programming solution and our implementation.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%Describe the background for chosen area that is going to be investigated. Write a short description of the area that is going to be %investigated. It is a brief description of the necessary background knowledge of the problem area and for carrying out the project.

In telecommunication one of the most challenging problems is the efficient allocation of the available frequency. Since the available bandwidth is always limited (and expensive), the efficient utilization of the frequency spectrum is a major concern. Due to the fastest growing number of mobile Internet users, the optimal channel allocation in cellular networks and their variants have been heavily researched in recent years \cite{Audhya:2011:SCA:1988563.1988571}.

Several variants of the channel allocation problem (CAP) have been defined based on the different channel constraints that a particular service might require. One of them is the so-called co-channel constraint where the same channel is not allowed to be assigned to neighboring cells simultaneously. This problem have been formalized as a graph coloring problem by many authors \cite{1456167}. Unfortunately, graph coloring is a well-known NP-complete problem \cite{Kar72} and therefore we do not know if a polynomial time algorithm for co-channel constraint satisfaction exists. Therefore various heuristic algorithms have been developed, the list of methods includes genetic algorithms, neural networks, graph-based and other approaches as well \cite{Audhya:2011:SCA:1988563.1988571}.

Cellular network topologies usually admit certain geometric structure. The most common network topology is the hexagonal grid topology where each cell is represented by a regular hexagon (two cells are neighbors if they share a common boundary). In \cite{662943}, the authors exploited this special structure and proposed an algorithm that optimally solves the CAP in $k$-band buffering systems where $k=1,2$. Moreover, the algorithm has polynomial running time $O(p)$ where $p$ is the number of cells.

The authors of \cite{7248845} introduced a \textit{distinctly different} CAP from all the above mentioned problems. Assuming a $2$-band buffering hexagonal cell topology (the interference graph created from this topology is called cellular graph) where each cell has a fixed number of frequency channel (channels are either busy or free), they asked the following question: "\textit{What is smallest size of the set of free channels associated with the cells (nodes of the cellular graph) that can guarantee interference free channel assignment to all the nodes?}". This problem is related to one of the generalizations of the graph coloring problem, called list coloring. It turned out that the required number of free channels is between $8$ and $10$. In addition, two algorithms have been proposed to create an interference free assignment, that is, a list-coloring of the cellular graph. The first one is the integer linear programming formulation of the list coloring problem (and therefore it is not a polynomial algorithm), the second one is a heuristic linear time algorithm that is, according to their experiment, within 12\% of the optimal solutions.

**add a short outline**

\section{Theoretical Framework}
In this section we introduce the basic notions and definitions that is essential to formulate the problem in a graph-theoretic setting.
\begin{defi} A graph $G$ is a  \textit{cellular graph} if it is constructed from the hexagonal cell topology in the following way: each cell is a node and two nodes are connected if and only if they share a common boundary.
\end{defi}
\begin{defi} A cellular network is $k$\textit{-band buffering} if the interference extends up to $k$ cells.
\end{defi}
\begin{defi} Let $G$ be a graph and $L(v)$ a set of colors for all $v \in V(G)$ such that $|L(v)|=k$. We say that $G$ is $k$\textit{-choosable} if $G$ is colorable such that the color of $v$ is in $L(v)$ for all $v \in V(G)$, such colorings called $k$\textit{-list coloring} of $G$.
\end{defi}
\begin{defi} The \textit{choice number} of $G$ is the smallest $k \in \mathbb{N}$ (notated by $\ch(G)$) such that $G$ is $k$-choosable.
\end{defi}

From now on we assume that all graphs are connected. We can do it without the loss of generality as all the algorithms and proofs can be repeated for each connected component of a graph.
\section{Problem formulation and research questions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acyclic orientations of cellular graphs}

% rewrite this
Let $G$ be a cellular graph of an $1$-band buffering cellular network. Find an orientation of $G$ such that the newly constructed directed graph $G'$ does not contain directed cycles such that $d^+(v) \leqslant 3$ ($d^+(v)$ is the indegree of $v$) for all $v \in V(G')$ in polynomial time of the size of $G'$. Construct a polynomial time algorithm that finds a kernel graph in $G'$, that is, an independent set $K \subseteq V(G)$ that satisfies the following: for each node $u \in V(G) \setminus K$ there is a node $v \in K$ such that $(u,v) \in E(G)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{lem}\label{lem:degree-constraint}
Let $G$ be a cellular graph. Then $G$ has a node $v$ such that $\deg(v) \leqslant 3$.
\end{lem}

\begin{proof} Let us consider the planar drawing of the graph $G$ which corresponds to its hexagonal topology.
Let $v$ be any node of $G$ and we assign $(0,0)$ to this node. We assign coordinates to every node starting from node $v$ in the following way. If a node $w$ is to the north of node $v$ then the coordinates of node $w$ is equal to the coordinates of node $v$ plus $(0,1)$. We summarize this method in Table \ref{table:assignment} according to the cardinal directions.
\begin{table}[!h]
\centering
 \begin{tabular}{||c|c||} 
 \hline
 Direction & Change  \\ [0.5ex] 
 \hline\hline
 N & $(0,1)$  \\ 
 \hline
 NE & $(1,1)$ \\
 \hline
 SE & $(1,0)$ \\
 \hline
 S & $(0,-1)$ \\
 \hline
 SW & $(-1,-1)$ \\
 \hline
 NW & $(-1,0)$ \\
 \hline
 \end{tabular}
  \caption{Rules of the hexagonal coordinate system}\label{table:assignment}
\end{table}
By applying this method to every node, we get a coordinate system within our hexagonal topology. Since there are finitely many node in $V(G)$, we can select the maximal node by selecting the node with maximal $x$- and $y$-coordinate, that is, a node $w$ with the following property: if $(x_1,y_1)$ is the coordinates of $w$ and $v \neq w$ is another node with the coordinates $(x_2,y_2)$ then $x_1 \geqslant x_2$ and $y_1 > y_2$ are satisfied.

To obtain a contradiction, suppose that the maximal node $w$ has more than 3 neighbors. It follows then at least one if its neighbors is either to the north, northeast or southeast to $w$. However, this neighbor would violate the maximality of $w$. Therefore we proved that $\deg(w) \leqslant 3$.
\end{proof}
\begin{defi}
We say that the orientation of a digraph $G$ is $k$\textit{-bounded} if for each node $v \in V(G)$ its outdegree is bounded by $k$, that is, $\deg^+(v) \leqslant k$.
\end{defi}
\begin{lem}\label{lem:bounded-acyclic-orientation}
If $G$ is a cellular graph, then $G$ has a $3$-bounded acyclic orientation.
\end{lem}
\begin{proof} Since $G=(V,E)$ is a cellular graph, we can make use of Lemma \ref{lem:degree-constraint}, that is, let $v$ be a node of $G^{(0)}:=G$ with $\deg(v) \leqslant 3$. We construct a function $f\colon V(G) \to \mathbb{N}$ by setting $f(v) := 0$ and then remove the node $v$ from $G^{(0)}$ yielding the graph $G^{(1)}$. By repeating the same procedure - at step $i$th - we select a node $v$ in $G^{(i)}$ such that $\deg(v) \leqslant 3$ and we set $f(v):=i$. Then we remove $v$ from $G^{(i)}$ which yields the graph $V(G^{(i+1)}) := V(G^{(i)}) \setminus \lbrace v \rbrace$. It is trivial that the graphs $G^{(1)},\ldots,G^{(|V(G)|)}$ are all cellular therefore it was valid using Lemma \ref{lem:degree-constraint}.

We note that the procedure ends after $|V(G)|$ steps.

Now we construct a digraph $H=(V,A)$ from $G$ using the function $f$. Let $(u,v) \in E$ be an arbitrary edge. If $f(u) < f(v)$ then $A:=A\cup (u,v)$ otherwise $A:=A \cup (v,u)$. By repeating this procedure for all edges in $E$ we get an orientation of $G$, that is, the digraph $H$. We need to prove that $H$ is
\begin{enumerate}
\item acyclic and
\item $\deg^+(v) \leqslant 3$ holds for all $v \in V(G)$.
\end{enumerate}
To prove (1), we will obtain a contradiction by assuming that $H$ contains a directed cycle. Let $C$ be a directed cycle in $H$ with the nodes $C:=\lbrace v_1,v_2,\ldots,v_k \rbrace$ that are ordered such that $(v_i,v_{i+1 \bmod{k}}) \in E$ $(i=1,2,\ldots,k)$ which means that $f(v_1) < f(v_2) < \ldots < f(v_{k-1}) < f(v_k)$ but also $f(v_k) < f(v_1)$ which is impossible.

What is left to show is that $\deg^+(v) \leqslant 3$. At each step we select a node with no more than $3$ neighbors and those neighbors will be assigned a number that is greater than $f(v)$. Therefore the outdegree of $v$ cannot be larger than $3$.
\end{proof}

It is easy to see that the proof of this lemma can be transformed into a polynomial time algorithm. In what follows we outline the algorithm (Algorithm \ref{alg:find-acyclic-bounded-orientation}) that is based on the proof of Lemma \ref{lem:bounded-acyclic-orientation} and then we prove that its running time is $O(|V(G)| + |E(G)|)$.
\begin{algorithm}\label{alg:find-acyclic-bounded-orientation}
 \KwData{A cellular graph $G = (V,E)$}
 \KwResult{An acyclic $3$-bounded orientation of $G$}
 $G^{(0)} := G$\;
 Initialize $f \colon V(G) \to \mathbb{N}$\;
 \For{$i\leftarrow 0$ \KwTo $|V(G)|-1$}{
  Select $v \in V(G^i)$ such that $\deg(v) \leqslant 3$\;
  $f(v):=i$\;
  $V(G^{(i+1)}):=V(G^{i}) \setminus \lbrace v \rbrace$\;
 }
 Initialize $H:=(V,A)$\;
 \ForAll{$e = (u,v) \in E(G)$}{
   \eIf{$f(u) < f(v)$} {
   	$A := A \cup (u,v)$\;
   }{
   $A := A \cup (v,u)$\;
   }   
 }
 \caption{Constructing an acyclic $3$-bounded orientation of a cellular graph}
\end{algorithm}
\begin{prop} Algorithm \ref{alg:find-acyclic-bounded-orientation} has a running time of $O(|V(G)| + |E(G)|)$.
\end{prop}
\begin{proof}
It is enough to prove that "Select $v \in V(G^i)$ such that $\deg(v) \leqslant 3$" can be done in $O(|V(G)|)$ since the rest is obvious. Let us initialize a queue $Q := \lbrace v \mid \deg(v) \leqslant  3\rbrace$ before we start running the algorithm. We pop a node $v$ from $G$ at every iteration and push new ones after the removal of $v$ if there are new nodes in $G$ that satisfy the degree condition. The pop and push methods can be implemented in constant time which completes the proof.
\end{proof}

It is important to note that the idea of Algorithm \ref{alg:find-acyclic-bounded-orientation} came from \cite{CHROBAK1991243} where the authors proved that every planar graph has a $5$-bounded acyclic orientation. They used the fact that every planar graph has a node with at most $5$ neighbors which follows from Euler's formula.
\section{Cellular graphs are $4$-choosable}

We are about to prove one of our main results, that is, cellular graphs are $4$-choosable. We would like to note that since $G$ is a planar, it follows from Thomassen's theorem \cite{Thomassen:1994:PG:184180.184192} that it is $5$-choosable. It is worth noting that cellular graphs are $3$-colorable \cite{662943} (Theorem 1). However, even with this additional constraint, we still cannot say that if $G$ is planar and $3$-colorable then it is $4$-choosable as a counterexample can be found in \cite{JGT:JGT4}. The counterexample is quite interesting (second figure) as it is constructed from a cellular graph however, the author added one extra node and joined it to all of the outer nodes. Due to this step, the resulting graph does not always contain a node of degree $3$ therefore the property of Lemma \ref{lem:degree-constraint} is not true for this construction.

Before proving our theorem we need to introduce the notion of kernel in graphs and a theorem from Galvin \cite{Galvin:1995:LCI:199352.199369} (non-multigraph version from \cite{citeulike:395714}, Lemma 5.4.3) that will play a central role in our study. The proof of the theorem is inductive and can be transformed into an algorithm.
\begin{defi} Let $G$ be a directed graph. We say that an independent set $K \subseteq V(G)$ is a kernel of $G$ if it satisfies the following: for each node $u \in V(G) \setminus K$ there is a node $v \in K$ such that $(u,v) \in E(G)$.
\end{defi}
\begin{thm}[Galvin, 1995]\label{thm:galvin} Let $G$ be a graph and $\lbrace L(v) : v \in V(G) \rbrace$ given color sets. If $G$ has an orientation $D$ such that $d^+(v) < |L(v)|$ for all $v \in V(D)$ and every induced subgraph of $D$ has a kernel, then $G$ can be colored from the given color sets. $\blacksquare$
\end{thm}
We will prove this theorem later through our algorithm. 
\begin{defi} We note that a directed graph is called kernel-perfect if all of its induced subgraphs contain a kernel.
\end{defi}
\begin{thm}\label{thm:choice-number}
If $G$ is a cellular graph then $\ch(G) \leqslant 4$.
\end{thm}
\begin{proof}
Since $G$ is cellular, we can consider its $3$-bounded acyclic orientation by Lemma \ref{lem:bounded-acyclic-orientation}. This orientation is kernel-perfect by Richardson's theorem \cite{richardson1946}. Therefore we can apply Theorem \ref{thm:galvin} (with $d^+(v) \leqslant 3$ and $L(v)=4$) which concludes the proof.
\end{proof}
We proved that cellular graphs are $4$-choosable what is left to show is that there exists an algorithm that constructs such colorings in polynomial time.
\section{Finding a kernel in DAGs}

It has been shown that finding a kernel in an arbitrary graph is $\mathrm{NP}$-complete \cite{chvatal}. However we will show that in DAGs (directed acyclic graphs) this problem can be easily solved in polynomial time. This is crucial as our $4$-list coloring algorithm has to construct kernels repeatedly during its running time.

\begin{lem} Let $G$ be a directed acyclic graph. Then a kernel $K \subseteq V(G)$ can be found in polynomial time.
\end{lem}
\begin{proof} Since $G$ is acyclic we can consider the topological order of its nodes (which can be computed in polynomial time). Let $\lbrace v_1, v_2, \ldots, v_n \rbrace$ be this topological order. Let us start from node $v_1$ and add it to the kernel, that is, $K = \lbrace v_1 \rbrace$. Mark the neighbors of $v_1$. After that, we move on with the second node $v_2$, if it is marked we skip it otherwise we add it to the kernel $(K := K \cup \lbrace v_2 \rbrace)$ and mark its neighbors. Repeat this procedure with the remaining nodes. We need to prove that
\begin{enumerate}
\item $K$ is independent and
\item for each node $u \in V(G) \setminus K$ there is a node $v \in K$ such that $(u,v) \in E(G)$.
\end{enumerate}
1) If $v \in K$ then its neighbors are marked. Since we do not add marked nodes to $K$, it is always independent.
2) When $v$ is unmarked and we add it to $K$ all of its neighbors come after it in the topological order (otherwise $v$ would have already been marked), that is, $v$ is connected to its neighbors by forward edges.

\begin{algorithm}\label{alg:find-kernel-in-dags}
 \KwData{A directed acyclic graph $G = (V,A)$}
 \KwResult{A kernel $K \subseteq V(G)$}
 $T := \text{the topological order of $V(G)$}$\;
 $K := \emptyset$\;
 \For{$i\leftarrow 1$ \KwTo $|V(G)|$}{
  \If{$T(i)$ is unmarked} {
  	$K := K \cup \lbrace T(i) \rbrace$\;
  	Mark the neighbors of $T(i)$\;
  }
 }
 \caption{Finding a kernel in $G$}
\end{algorithm}
We note that the topological order can be computed in linear time in the number of nodes and edges with Kahn's algorithm \cite{Kahn:1962:TSL:368996.369025} therefore the running time of our algorithm is $O(|V(G)|+|E(G)|)$.
\end{proof}
\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{IEEEtran}
\bibliography{ref.bib}
\end{document}